---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the .Rmd file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# FFTrees `r packageVersion("FFTrees")`

<!-- Status badges: --> 

[![Build Status](https://travis-ci.org/ndphillips/FFTrees.svg?branch=master)](https://travis-ci.org/ndphillips/FFTrees)
[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/FFTrees)](https://CRAN.R-project.org/package=FFTrees)
[![Downloads](https://cranlogs.r-pkg.org/badges/FFTrees?color=brightgreen)](https://www.r-pkg.org/pkg/FFTrees)

<!-- Goal of R package: -->

The R package **FFTrees** creates, visualizes and evaluates _fast-and-frugal decision trees_ (FFTs) from data with a binary outcome following the methods described in Phillips, Neth, Woike & Gaissmaier (2017, [html](https://journal.sjdm.org/17/17217/jdm17217.html) | [PDF](https://journal.sjdm.org/17/17217/jdm17217.pdf)). 


## Installation

You can install the released version of **FFTrees** [from CRAN](https://CRAN.R-project.org/package=FFTrees) with:

```{r install-cran, eval = FALSE}
install.packages("FFTrees")
```

And the development version [from GitHub](https://github.com/ndphillips/FFTrees) with:

```{r install-gitub, eval = FALSE}
# install.packages("devtools")
devtools::install_github("ndphillips/FFTrees", build_vignettes = TRUE)
```

## Purpose 

_Fast-and-frugal trees_\ (FFTs) are decision algorithms for solving binary classification tasks in an effective and efficient fashion. 
Their performance when predicting cases for new data competes with machine learning techniques such as logistic regression\ (LR), support-vector machines\ (SVM), and random forests\ (RF), but are typically simpler and more transparent. 

## Examples

As an example, let's create a FFT predicting heart disease status (_Healthy_ vs. _Diseased_) based on the `heartdisease` dataset included in **FFTrees**: 

```{r load-pkg}
library(FFTrees)

dim(heartdisease)
```

### Using data 

The `heartdisease` data provides medical information for 303\ patients that were tested for heart disease. 
The full data were split into two subsets: 
A `heart.train` dataset for fitting decision trees, and `heart.test` dataset for a testing the resulting trees. 
Here are the first rows and columns of both subsets of the `heartdisease` data: 

- `heart.train` (the training / fitting dataset) contains the data from `r nrow(heart.train)` patients: 

```{r data-train}
heart.train
```

- `heart.test` (the testing / prediction dataset) contains data from a new set of `r nrow(heart.test)` patients:

```{r data-test}
heart.test
```

The criterion variable is `diagnosis`, a logical column  (either\ `TRUE` or\ `FALSE`) which indicate the true state for each patient (i.e., whether or not they suffer from heart disease). 

### Creating fast-and-frugal trees (FFTs) 

Now let's use `FFTrees()` to create fast-and-frugal trees from the `heart.train` data and test their performance on the `heart.test` data: 

```{r set-seed, echo = FALSE}
set.seed(246)  # for reproducible randomness
```

```{r example-heart, fig.width = 6, fig.height = 6, out.width = "80%", fig.align='center'}
# Create an FFTrees object from the heartdisease data: 
heart.fft <- FFTrees(formula = diagnosis ~., 
                     data = heart.train,
                     data.test = heart.test, 
                     decision.labels = c("Healthy", "Disease"))

# Printing an FFTrees object shows aggregate statistics: 
heart.fft

# Plot the best tree applied to the test data: 
plot(heart.fft,
     data = "test",
     main = "Heart Disease")

# Compare results across algorithms in test data: 
heart.fft$competition$test
```

<!-- FFTs by verbal description: -->

### Building FFTs from verbal description 

Because fast-and-frugal trees are so simple, we even can create one 'from words' and apply it to data!

For example, let's create a tree with the following three nodes and evaluate its performance on the `heart.test` data:

1. If `sex = 1`, predict _Disease_.
2. If `age < 45`, predict _Healthy_.
3. If `thal = {fd, normal}`, predict _Healthy_,   
otherwise, predict _Disease_. 

These conditions can directly be supplied to the `my.tree` argument of `FFTrees()`: 

```{r example-heart-verbal, fig.width = 6, fig.height = 6, out.width = "80%", fig.align='center'}
# Create custom FFT 'in words' and apply it to data:

# Create my own FFT (from verbal description):
my.fft <- FFTrees(formula = diagnosis ~., 
                  data = heart.train,
                  data.test = heart.test, 
                  decision.labels = c("Healthy", "Disease"),
                  my.tree = "If sex = 1, predict Disease.
                             If age < 45, predict Healthy.
                             If thal = {fd, normal}, predict Healthy,  
                             Otherwise, predict Disease.")

# Plot and evaluate my custom FFT:
plot(my.fft,
     data = "test",
     main = "My custom FFT")
```

As we can see, this particular tree is somewhat biased: 
It has nearly perfect _sensitivity_ (i.e., good at identifying cases of _Disease_) but suffers from low _specificity_ (i.e., not so good at identifying _Healthy_ cases).  Overall, the _accuracy_ of our custom tree exceeds the data's baseline by a fair amount, but exploring the **FFTrees** package further will make it easy to design better FFTs.

## Citation

We had a lot of fun creating **FFTrees** and hope you like it too! 
We have an article introducing the **FFTrees**  package in the journal _Judgment and Decision Making_, entitled [FFTrees: A toolbox to create, visualize,and evaluate fast-and-frugal decision trees](https://journal.sjdm.org/17/17217/jdm17217.pdf) (\ [html](https://journal.sjdm.org/17/17217/jdm17217.html) | [PDF](https://journal.sjdm.org/17/17217/jdm17217.pdf)\ ).  


**Citation** (in APA format): 

- Phillips, N. D., Neth, H., Woike, J. K. & Gaissmaier, W. (2017). 
FFTrees: A toolbox to create, visualize, and evaluate fast-and-frugal decision trees. 
_Judgment and Decision Making_, _12_ (4), 344–368. 


We encourage you to read the article to learn more about the history of FFTs and how the FFTrees package creates them. 
If you use **FFTrees** in your work, please cite us and spread the word so we can continue developing the package. 

<!-- Examples uses/publications: -->

Here are some example publications that have used **FFTrees** (find the full list at [Google Scholar](https://scholar.google.com/scholar?oi=bibs&hl=en&cites=205528310591558601)): 

- [Lötsch, J., Haehner, A., & Hummel, T. (2020). Machine-learning-derived rules set excludes risk of Parkinson’s disease in patients with olfactory or gustatory symptoms with high accuracy. 
_Journal of Neurology_, _267_(2), 469-478.](https://link.springer.com/article/10.1007/s00415-019-09604-6)

- [Kagan, R., Parlee, L., Beckett, B., Hayden, J. B., Gundle, K. R., & Doung, Y. C. (2020). Radiographic parameter-driven decision tree reliably predicts aseptic mechanical failure of compressive osseointegration fixation. _Acta Orthopaedica_, _91_(2), 171-176.](https://www.tandfonline.com/doi/full/10.1080/17453674.2020.1716295)

- [Klement, R. J., Sonke, J. J., Allgäuer, M., Andratschke, N., Appold, S., Belderbos, J., ... & Mantel, F. (2020). Correlating dose variables with local tumor control in stereotactic body radiotherapy for early stage non-small cell lung cancer: A modelling study on 1500 individual treatments. 
_International Journal of Radiation Oncology * Biology * Physics_.](https://www.sciencedirect.com/science/article/pii/S036030162030897X)

- [Nobre, G. G., Hunink, J. E., Baruth, B., Aerts, J. C., & Ward, P. J. (2019). 
Translating large-scale climate variability into crop production forecast in Europe. 
_Scientific Reports_, _9_(1), 1-13.](https://www.nature.com/articles/s41598-018-38091-4)

- [Buchinsky, F. J., Valentino, W. L., Ruszkay, N., Powell, E., Derkay, C. S., Seedat, R. Y., ... & Mortelliti, A. J. (2019). Age at diagnosis, but not HPV type, is strongly associated with clinical course in recurrent respiratory papillomatosis. _PloS One_, _14_(6).](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6563955/)


<!-- footer: -->

----

[File `README.Rmd` last updated on `r Sys.Date()`.] 

<!-- eof. -->
