---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# FFTrees

[![Build Status](https://travis-ci.org/ndphillips/FFTrees.svg?branch=master)](https://travis-ci.org/ndphillips/FFTrees)
[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/FFTrees)](https://CRAN.R-project.org/package=FFTrees)
[![Downloads](https://cranlogs.r-pkg.org/badges/FFTrees?color=brightgreen)](https://www.r-pkg.org/pkg/FFTrees)

The goal of FFTrees is to create and visualize fast-and-frugal decision trees (FFTs) from data with a binary outcome following the methods described in [Phillips, Neth, Woike & Gaissmaier (2017)](http://journal.sjdm.org/17/17217/jdm17217.pdf).

## Installation

You can install the released version of FFTrees from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("FFTrees")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("ndphillips/FFTrees", build_vignettes = TRUE)
```
## Examples

```{r}
library(FFTrees)
```

Let's create a fast-and-frugal tree predicting heart disease status ("Healthy" vs. "Diseased") based on a `heart.train` dataset, and test it on `heart.test` a testing dataset.

Here are the first new rows and columns of our datasets. The key column is `diagnosis`, a logical column  (TRUE and FALSE) which indicate, for each patient, whether or not they have heart disease.

Here is `heart.train` (the training dataset) which contains data from `r nrow(heart.train)` patients:

```{r}
heart.train
```

Here is `heart.test` (the testing / prediction dataset)  which contains data from a new set of `r nrow(heart.test)` patients:

```{r}
heart.test
```

Now let's use `FFTrees()` to create a fast and frugal tree from the `heart.train` data and test their performance on `heart.test`

```{r example, fig.width = 6, fig.height = 6, out.width = "80%"}
# Create an FFTrees object from the heartdisease data
heart.fft <- FFTrees(formula = diagnosis ~., 
                     data = heart.train,
                     data.test = heart.test, 
                     decision.labels = c("Healthy", "Disease"))

# See the print method which shows aggregatge statistics
heart.fft

# Plot the best tree applied to the test data
plot(heart.fft,
     data = "test",
     main = "Heart Disease")

# Compare results across algorithms in test data
heart.fft$competition$test
```

Because fast-and-frugal trees are so simple, you can create one 'from words' and apply it to data!

```{r fig.width = 6, fig.height = 6, out.width = "80%"}
# Create your own custom FFT 'in words' and apply it to data

# Create my own fft
my.fft <- FFTrees(formula = diagnosis ~., 
                  data = heart.train,
                  data.test = heart.test, 
                  decision.labels = c("Healthy", "Disease"),
                  my.tree = "If sex = 1, predict Disease.
                             If age < 45, predict Healthy.
                             If thal = {fd, normal}, predict Disease. 
                             Otherwise, predict Healthy")

# Plot my custom fft and see how it did
plot(my.fft,
     data = "test",
     main = "Custom FFT")
```


## Citation

*APA Citation*

Phillips, Nathaniel D., Neth, Hansjoerg, Woike, Jan K., & Gaissmaier, W. (2017). FFTrees: A toolbox to create, visualize, and evaluate fast-and-frugal decision trees. *Judgment and Decision Making*, 12(4), 344-368.

We had a lot of fun creating FFTrees and hope you like it too! We have an article introducing the `FFTrees` package in the journal Judgment and Decision Making titled FFTrees: [A toolbox to create, visualize,and evaluate fast-and-frugal decision trees](http://journal.sjdm.org/17/17217/jdm17217.pdf). We encourage you to read the article to learn more about the history of FFTs and how the FFTrees package creates them.

If you use FFTrees in your work, please cite us and spread the word so we can continue developing the package

Here are some example publications that have used FFTrees (find the full list at [Google Scholar](https://scholar.google.com/scholar?oi=bibs&hl=en&cites=205528310591558601))

- [Lötsch, J., Haehner, A., & Hummel, T. (2020). Machine-learning-derived rules set excludes risk of Parkinson’s disease in patients with olfactory or gustatory symptoms with high accuracy. Journal of Neurology, 267(2), 469-478.](https://link.springer.com/article/10.1007/s00415-019-09604-6)
- [Kagan, R., Parlee, L., Beckett, B., Hayden, J. B., Gundle, K. R., & Doung, Y. C. (2020). Radiographic parameter-driven decision tree reliably predicts aseptic mechanical failure of compressive osseointegration fixation. Acta Orthopaedica, 91(2), 171-176.](https://www.tandfonline.com/doi/full/10.1080/17453674.2020.1716295)
- [Klement, R. J., Sonke, J. J., Allgäuer, M., Andratschke, N., Appold, S., Belderbos, J., ... & Mantel, F. (2020). Correlating dose variables with local tumor control in stereotactic body radiotherapy for early stage non-small cell lung cancer: A modelling study on 1500 individual treatments. International Journal of Radiation Oncology* Biology* Physics.](https://www.sciencedirect.com/science/article/pii/S036030162030897X)
- [Nobre, G. G., Hunink, J. E., Baruth, B., Aerts, J. C., & Ward, P. J. (2019). Translating large-scale climate variability into crop production forecast in Europe. Scientific reports, 9(1), 1-13.](https://www.nature.com/articles/s41598-018-38091-4)
- [Buchinsky, F. J., Valentino, W. L., Ruszkay, N., Powell, E., Derkay, C. S., Seedat, R. Y., ... & Mortelliti, A. J. (2019). Age at diagnosis, but not HPV type, is strongly associated with clinical course in recurrent respiratory papillomatosis. PloS one, 14(6).](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6563955/)

