---
title: "Example FFTs"
author: "Nathaniel Phillips"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Examples of FFTs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, echo = F, message = F, results = 'hide'}
library(FFTrees)
```

## Mushrooms

The `mushrooms` dataset contains data about mushrooms. The goal of our model is to predict which mushrooms are poisonous.

First, we'll create a training and test dataset:

```{r}
set.seed(100)
train.cases <- sample(c(T, F), size = nrow(mushrooms), replace = T)
mushrooms.train <- mushrooms[train.cases,]
mushrooms.test <- mushrooms[train.cases == F,]
```


```{r}
mushrooms.fft <- fft(formula = poisonous ~.,
                     data = mushrooms.train,
                     data.test = mushrooms.test)
```

Here is a summary of the final 8 trees:

```{r}
summary(mushrooms.fft)
```

Here is the best training tree applied to the test data:

```{r, fig.width = 6, fig.height = 6}
plot(mushrooms.fft, data = "test")
```


## Iris

The `iris` dataset contains data about 150 flowers. Our goal is to predict which flowers are of the class Virginica

```{r}
head(iris)
```

We'll create a new fft object called `iris.fft` using the `fft()` function. In this case, we won't create a separate training and test dataset, we'll just fit the tree to the entire dataset:

```{r}
iris.fft <- fft(
  formula = virginica ~.,
  data = iris
  )
```

Here is the final best training tree (tree 6):

```{r, fig.width = 6, fig.height = 6}
plot(iris.fft)
```

Here is tree number 2, which is a bit more conservative than tree 6.

```{r, fig.width = 6, fig.height = 6}
plot(iris.fft, which.tree = 2)
```
