---
title: "Overview of FFTrees"
author: "Nathaniel Phillips"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: fft.bib
csl: apa.csl
vignette: >
  %\VignetteIndexEntry{Overview FFTrees}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## What is a Fast and Frugal Decision Tree (FFT)?

Informally, a fast and frugal decision tree (FFT) is a set of rules for making decisions based on very little information (usually 1 to 5 pieces). Formally, a FFT is a decision tree has at least one *exit* on each level of the tree [@martignon2003naive]. FFTs are simple, transparent decision strategies that use minimal information to make decisions [see @gigerenzer1999fast;@gigerenzer1999good]. They are frequently prefereable to more complex decision strategies (such as Logistic Regression) because they rarely overfit data [@gigerenzer2009homo] and are easy to interpret and impliment in real-world decision tasks [@marewski2012heuristic]. They have been used in real world tasks from detecting depression [@jenny2013simple], to making fast decisions in emergency rooms [@green1997alters].

The purpose of the `FFTrees` package is to produce, compare, and display FFTs like the one below. 

```{r, echo = F, message = F, results = 'hide'}
library(FFTrees)
```


```{r, fig.width = 6, fig.height = 6, echo = T, fig.align='center'}
bcancer.fft <- fft(diagnosis ~.,
                   data = breastcancer
                 )

plot(bcancer.fft,
     description = "Breast Cancer",
     decision.names = c("Absent", "Present"))
```

The main function in the package is `fft()` which takes a formula `formula` and a training dataset `data` arguments and returns several FFTs which attempt to classify training cases into criterion classes. The `fft()` function returns a list object with the "fft" class which can then be passed to other functions such as `plot()` (which plots the FFTs), and `predict()` which applies an existing set of FFTs to new datasets.

## Datasets

The package contains several datasets taken from the [UCI Machine Learning Repository](http://archive.ics.uci.edu/ml/).

- `heartdisease` -- patients suspected of having heart disease [source](http://archive.ics.uci.edu/ml/datasets/Heart+Disease)
- `breastcancer` -- patients suspected of having breast cancer [source](https://archive.ics.uci.edu/ml/datasets/Breast+Cancer+Wisconsin+(Diagnostic))
- `titanic` -- records of which passengers on the Titanic survived
- `forestfires` -- forest fire statistics [source](http://archive.ics.uci.edu/ml/datasets/Forest+Fires)
- `wine` -- ratings of wine quality [source](http://archive.ics.uci.edu/ml/datasets/Wine)
- `income` -- Census data from > 30,000 US residents [source](http://archive.ics.uci.edu/ml/datasets/Adult)
- `bank` -- Bank marketing dataset [source](http://archive.ics.uci.edu/ml/datasets/Bank+Marketing)

## Guides

To learn more about the package, click the following guides:

- [Creating FFTs with fft()](fft_function.html)
- [Visualizing FFTs with plot()](fft_plot.html)
- [Lots of example FFTs](fft_examples.html)

## Bibliography
