---
title: "Examples of FFTs"
author: "Nathaniel Phillips"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Examples of FFTs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, echo = F, message = F, results = 'hide'}
library(FFTrees)
```

# Examples

Here are some examples of FFTs created from datasets in the `FFTrees` package

## Mushrooms

The `mushrooms` dataset contains data about mushrooms (see `?mushrooms` for details). The goal of our model is to predict which mushrooms are poisonous based on `r ncol(mushrooms) - 1` cues ranging from the mushroom's odor, color, etc. 

Let's create some ffts. We'll separate the data into a 50% training and a 50% testing dataset with the `train.p = .5` argument:

```{r}
set.seed(100) # For replicability of the training / test data split

mushrooms.fft <- fft(formula = poisonous ~.,
                     data = mushrooms,
                     train.p = .5)
```

Here's basic information about the trees:

```{r}
mushrooms.fft
```

[Cool beans.](https://goo.gl/B7YDuC)

### mushrooms cue accuracies

First, let's look at the individual cue training accuracies with `showcues`:

```{r fig.width = 6, fig.height = 6, fig.align = 'center'}
showcues(mushrooms.fft, main = "Mushrooms")
```

It looks like the cues *oder* and *sporepc* are the best predictors. in fact, the single cue *odor* has a hit rate of 97% and a false alarm rate of 0%! Based on this, we should expect the final FFT to have no more than these two cues:

### mushrooms FFT

Now let's plot the best tree applied to the test dataset

```{r, fig.width = 6, fig.height = 6, fig.align = 'center'}
plot(mushrooms.fft, 
     data = "test", 
     description = "Mushrooms FFT",
     decision.names = c("Safe", "Poisonous")
     )
```

Indeed, it looks like the best tree only uses the *odor* and *sporepc* cues. In our test dataset, the tree had a false alarm rate of 0% (1 - specificity), a hit rate of 85%, and a d-prime of 2.27.

Now, let's say that you talk to a mushroom expert who says that we are using the wrong cues. According to her, the best predictors for poisonous mushrooms are *ringtype* and *ringnum*. Let's build a set of trees with these cues and see how they perform relative to our initial tree:

```{r}
set.seed(100) # for replicability
mushrooms.ring.fft <- fft(formula = poisonous ~ ringtype + ringnum,
                          data = mushrooms,
                          train.p = .5
                          )
```

Here is the final tree:

```{r, fig.width = 6, fig.height = 6, fig.align = 'center'}
plot(mushrooms.ring.fft, 
     data = "test", 
     description = "Mushrooms (Ring only) FFT",
     decision.names = c("Safe", "Poisonous")
     )
```

As we can see, this tree did not perform nearly as well as our earlier one. The new d-prime is only 0.84. 


## Iris

The `iris` dataset contains data about 150 flowers (see `?iris`). Our goal is to predict which flowers are of the class Virginica

```{r}
iris.fft <- fft(
  formula = virginica ~.,
  data = iris,
  train.p = .5
  )
```

### iris cue accuracies

First, let's look at the individual cue training accuracies with `showcues`:

```{r fig.width = 6, fig.height = 6, fig.align = 'center'}
showcues(iris.fft)
```

It looks like the cues *pet.wid* and *pet.len* are the best predictors. Based on this, we should expect the final FFT will likely use just one or both of these cues

### iris FFT

Now let's plot the best tree applied to the test dataset

```{r, fig.width = 6, fig.height = 6, fig.align = 'center'}
plot(iris.fft, 
     data = "test", 
     description = "Iris FFT",
     decision.names = c("Not V", "Virginica")
     )
```

Indeed, it looks like the best tree only uses the *pet.wid* and *pet.len* cues. In our test dataset, the tree had a false alarm rate of 2% (1 - specificity), a hit rate of 100%, and a d-prime of 2.15.

Now, this tree did quite well, but what if someone wants a tree with the maximum possible hit-rate. If we look at the ROC plot in the bottom right corder of the plot above, we can see that tree #6 has a hit-rate of 100%. Let's look at that tree:

```{r, fig.width = 6, fig.height = 6, fig.align = 'center'}
plot(iris.fft, 
     data = "test", 
     description = "Iris FFT",
     decision.names = c("Not V", "Virginica"),
     which.tree = 6     # Show tree #6
     )
```

As you can see, this tree does indeed have a maximum hit-rate. However, it comes at a high cost to specificity. 

