---
title: "Manually specifying FFTs"
author: "Nathaniel Phillips"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: fft.bib
csl: apa.csl 
vignette: >
  %\VignetteIndexEntry{Manually specifying FFTs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      fig.width = 7.5, 
                      fig.height = 7.5, 
                      dpi = 100, 
                      out.width = "600px", 
                      fig.align='center', 
                      message = FALSE)
```


```{r load-pkg-0, echo = FALSE, message = FALSE, results = 'hide'}
library(FFTrees)
```

## Manually specifying FFTs 

We typically create fast-and-frugal trees (FFTs) from data by using the `FFTrees()` function (see the [Main guide](guide.html), the vignette on [Creating FFTs with FFTrees()](FFTrees_function.html) and for details). 
However, we sometimes may want to design and test some specific FFT

There are two ways to define fast-and-frugal trees manually when using the `FFTrees()` function: 

1. as a sentence using the `my.tree` argument (the easier way), or  
2. as a data frame using the `tree.definitions` argument (the harder way).  

Both of these methods will bypass the tree construction algorithms built into the **FFTrees** package.


### `my.tree`

The first method is to use the `my.tree` argument, where `my.tree` is a sentence describing a (single) FFT. When this argument is specified in `FFTrees()`, the function (specifically `wordstoFFT()` will try to extract the specified FFT from the argument.

For example, let's look at the columns sex, age and thal in the `heartdisease` data:

```{r data}
head(heartdisease[c("sex", "age", "thal")])
```

Here's how we could specify an FFT using these cues as a sentence:

```{r fft-description, eval = FALSE}
my.tree = "If sex = 1, predict True.
           If age < 45, predict False. 
           If thal = {fd, normal}, predict True. Otherwise, predict False"
```

Here are some notes on specifying trees manually:

- Each node must start with the word "If" and should in the form: `If <CUE> <DIRECTION> <THRESHOLD>, predict <EXIT>`.
- Numeric thresholds shold be specified directly (without brackets).
- Factor thresholds must be specified within braces like `sex = {male}`. For factors with sets of values, values within a threshold should be separated by commas like `eyecolor = {blue,brown}`
- Standard logical comparisons `=`, `!=`, `<`, `>=` (etc.) are valid. For numeric cues, only use `>`, `>=`, `<`, `<=`. For factors, only use `=` and `!=`.
- Positive exits are indicated by `True`, while negative exits are specified by `False`. The final node will be forced to have a bidirectional exit. The text `Otherwise, predict EXIT` I've included in the example above is actually not necessary.


Now, let's pass the `my.tree` argument to `FFTrees()` to force apply our FFT to the heartdisease data:

```{r my-tree}
# Pass a verbally defined FFT to FFTrees with the my.tree argument
my.heart.fft <- FFTrees(diagnosis ~.,
                        data = heartdisease,
                        my.tree = "If sex = 1, predict True.
                                   If age < 45, predict False. 
                                   If thal = {fd, normal}, predict True. 
                                   Otherwise, predict False")
```

Let's see how well our FFT did:

```{r plot-my-tree}
# Plot 
plot(my.heart.fft)
```

As you can see, this FFT is pretty terrible --- it has a high sensitivity (of\ 85%), but much lower specificity (of\ 63%). 

Let's see if we can come up with a better FFT by using the cues\ `thal`, `cp`, and\ `ca`: 

```{r verbal-spec-fft}
# Specify an FFt verbally with the my.tree argument: 
my.heart.fft <- FFTrees(diagnosis ~.,
                        data = heartdisease,
                        my.tree = "If thal = {rd,fd}, predict True.
                                   If cp != {a}, predict False. 
                                   If ca > 1, predict True. 
                                   Otherwise, predict False")

# Plot resulting FFT:
plot(my.heart.fft)
```

This FFT looks much better --- but check out whether you can create even better ones! 


## Vignettes

<!-- Table of all vignettes: -->

Here is a complete list of the vignettes available in the **FFTrees** package: 

|   | Vignette | Description |
|--:|:------------------------------|:-------------------------------------------------|
|   | [Main guide](guide.html) | An overview of the **FFTrees** package |
| 1 | [Heart Disease Tutorial](FFTrees_heart.html)   | An example of using `FFTrees()` to model heart disease diagnosis |
| 2 | [Accuracy statistics](AccuracyStatistics.html) | Definitions of accuracy statistics used throughout the package |
| 3 | [Creating FFTs with FFTrees()](FFTrees_function.html) | Details on the main function `FFTrees()` |
| 4 | [Specifying FFTs directly](FFTrees_mytree.html)   | How to directly create FFTs with `my.tree` without using the built-in algorithms |
| 5 | [Visualizing FFTs with plot()](FFTrees_plot.html) | Plotting `FFTrees` objects, from full trees to icon arrays |
| 6 | [Examples of FFTs](FFTrees_examples.html) | Examples of FFTs from different datasets contained in the package |


<!-- eof. -->
