---
title: "FFTrees tree construction algorithms"
author: "Nathaniel Phillips"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: fft.bib
csl: apa.csl
vignette: >
  %\VignetteIndexEntry{FFTrees tree construction algorithms}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(FFTrees)
options(digits = 3)
```


FFTrees currently contains 4 different tree construction algorithms:

| Algorithm| Full Name| Reference
|:------|:----|:---|
|     ifan|    Marginal Fan| @phillips2017FFTrees  |
|     dfan|    Conditional Fan|@phillips2017FFTrees|  
|     zigzag| Zig-Zag  | @martignon2008categorization|
|     max| Max  |   @martignon2008categorization |


##"ifan"

The default algorithm used to create trees `"ifan"` can be summarised in five steps. 

|     Step| Function|Description |
|:-----|:----|:----------------------------------------------------|
|  1|    `cuerank`|For each cue, calculate a *classification threshold* that maximizes `goal.chase` (default is `bacc`) of classifications of *all data* based on that cue (that is, ignoring all other cues). If the cue is numeric, the threshold is a number. If the cue is a factor, the threshold is one or more factor levels.|
| 2|    `grow.FFTrees()`|Rank cues in order of their highest balanced accuracy value calculated using the classification threshold determined in step 1|
|  3|    `grow.FFTrees()`|Create all possible trees by varying the exit direction (left or right) at each level to a maximum of X levels (default of `max.levels = 4`).     |
|  4|    `grow.FFTrees()`|Reduce the size of trees by removing (pruning) lower levels containing less than X\% (default of `stopping.par = .10`) of the cases in the original data.|
|  5|    `grow.FFTrees()`|Select the FFT with the highest `goal` (default is `bacc` value as the final tree (`tree.max`)|

Table: 5 Steps in growing FFTs using the ifan algorithm.

## Example: Heart Disease

First, we'll calculate a classification threshold for each cue using `cuerank()`:

```{r, message = FALSE}
heartdisease.ca <- FFTrees::cuerank(formula = diagnosis ~., 
                                    data = heartdisease)

# Print key results
heartdisease.ca[c("cue", "threshold", "direction", "bacc")]
```

Here, we see the best decision threshold for each cue that maximizes its balanced accuracy (`bacc`) when applied to the entire dataset (independently of other cues). For example, for the age cue, the best threshold is `r with(heartdisease.ca, paste("age", direction[cue == "age"], threshold[cue == "age"]))` which leads to a balanced accuracy of 0.63. In other words, if we only had the age cue, then the best decision is: "If age > 55, predict heart disease, otherwise, predict no heart disease".

Let's confirm that this threshold makes sense. To do this, we can plot the bacc value for all possible thresholds as in Figure \@ref(fig:agethreshold):

```{r agethreshold, fig.width = 6, fig.height = 6, fig.align='center', echo = FALSE, fig.cap="Plotting the balanced accuracy of each decision threshold for the age cue."}
# Age thresholds to test
ages.to.test <- seq(20, to = 90, by = 5)

# Calculate balanced accuracy for each threshold
y <- sapply(ages.to.test, FUN = function(x) {
  
  decisions <- heartdisease$age > x
  
  sens <- mean(decisions[heartdisease$diagnosis == 1] == 1)
  spec <- mean(decisions[heartdisease$diagnosis == 0] == 0)
  
  bacc <- (sens + spec) / 2
  
  return(bacc)

})

plot(ages.to.test, y, 
     type = "b", 
     xlab = "age threshold", 
     ylab = "balanced accuracy (bacc)", ylim = c(.5, .7))

grid()

text(55, 
     y[ages.to.test == 55], 
     labels = paste("age = 55, bacc = ", round(y[ages.to.test == 55], 2)), 
     pos = 3)
```

Next, the cues are ranked by their balanced accuracy. Let's do that with the heart disease cues:

```{r}
# Rank heartdisease cues by balanced accuracy
heartdisease.ca <- heartdisease.ca[order(heartdisease.ca$bacc, decreasing = TRUE),]

# Print the key columns
heartdisease.ca[c("cue", "threshold", "direction", "bacc")]
```

Now, we can see that the top five cues are `thal`, `cp`, `ca`, `thalach` and `exang`. Because ffts rarely exceed 5 cues, we can expect that the trees will use a subset (not necessarily all) of these 5 cues.

We can also plot the cue accuracies in ROC space using the `showcues()` function:

```{r, fig.width = 7, fig.height = 7, fig.align = 'center', fig.cap = "Cue accuracies for the heartdisease dataset. The top 5 cues in terms of balanced accuracy are highlighted."}
# Show the accuracy of cues in ROC space
showcues(cue.accuracies = heartdisease.ca)
```

Next, `grow.FFTrees()` will grow several trees from these cues using different exit structures:

```{r}
# Grow FFTs
heartdisease.ffts <- grow.FFTrees(formula = diagnosis ~., 
                                  data = heartdisease)

# Print the tree definitions
heartdisease.ffts$tree.definitions
```

Here, we see that we have 7 different trees, each using some combination of the top 5 cues we identified earlier. For example, tree 1 uses the top 4 cues, while tree 3 uses only the top 3 cues. Why is that? The reason is that the algorithm also *prunes* lower branches of the tree if there are too few cases classified at lower levels. By default, the algorithm will remove any lower leves that classfify fewer than 10\% of the original cases. The pruning criteria can be controlled using the `stopping.rule`, `stopping.par` and `max.levels` arguments in `grow.FFTrees()`

Now let's use the wrapper function `FFTrees()` to create the trees all at once. We will then plot tree #4 which, according to our results above, should contain the cues `thal, cp, ca`"

```{r, fig.width = 5, fig.height = 5, message = FALSE, fig.align = "center"}
# Create trees
heart.fft <- FFTrees(formula = diagnosis ~., 
                     data = heartdisease)

# Plot tree # 4
plot(heart.fft, 
     stats = FALSE,    # Don't include statistics
     tree = 4)
```


## `"dfan"`

The `"dfan"` algorithm is identical to the `"ifan"` algorithm with one (important) exception: In `algorithm = "dfan"`, the thresholds and rankings of cues are recalculated for each level in the FFT conditioned on the exemplars that were *not* classified at higher leves in the tree. For example, in the `heartdisease` data, using `algorithm = "dfan"` would first classify some cases using the `thal` cue at the first level, and would then calculate new accuracies for the remaining cues on the remaining cases that were not yet classified. This algorithm is appropriate for datasets where cue validities systematically differ for different (and predictable) subsets of data. However, because it calculates cue thresholds for increasingly smaller samples of data as the tree grows, it is also, potentially, more prone to overfitting compared to `algorithm = "ifan"`


## Additional arguments

One can adjust the `"ifan"` and `"dfan"` algorithms in multiple ways. The most important arguments are `goal` and `goal.chase` which affect how thresholds are determined for each cue, and how cues are selected and ordered.

| Argument| Functionality| Default | Other possible arguments
|:------|:-------------------------|:-------|:---------|
|     `goal.chase`| Specifies which statistic is optimized when *calculating* cue thresholds and ranking cues. |`goal.chase = "bacc"`| `"wacc"`, `"acc"`, `"dprime"`, `"cost"` |
|     `goal`| Specifies which statistic is optimized when *selecting* an FFT in the fan.|`goal = "wacc"`| `"acc"`, `"dprime"`, `"cost"` |


## Specifying costs

If `goal = "cost"` and/or `goal.chase = "cost"`, the fan algorithms will try to minimize costs. One can specify two types of costs, `cost.cues`, the cost of using a cue to classify a case,, and `cost.outcomess` the cost of different outcomes.

`cost.cues` should be a data frame with two columns, one column giving the names of cues with costs, and one column giving the actual costs. For example, in the `heartdisease` dataset, cues have the following costs:

```{r}
heart.cue.cost <- data.frame("cue" = c("age", "sex", "cp", "trestbps", "chol",  "fbs", "restecg", "thalach", "exang", "oldpeak", "slope", "ca", "thal"),
                         "cost" = c(1, 1, 1, 1, 7.27, 5.2, 15.5, 102.9, 87.3, 87.3, 87.3, 100.9, 102.9))
```

`cost.outcome` should be a vector of length 4 indicating the cost of hits, false alarms, misses, and correct rejections reespectively. For example, if a false alarm has a cost of \$500 and a miss has a cost of $1000, one could specify:

```{r}
# Specify the following costs for heart disease diagnosis:
# cost(Hit) = 0, cost(False Alarm) = 100, cost(Miss) = 200, cost(correct rejection) = 0

heart.cost.outcomes <- c(0, 500, 1000, 0)
```


Here is an FFT that is built with the goal of maximizing balanced accuracy and *ignoring* these costs:

```{r}
heart.costA.fft <- FFTrees(formula = diagnosis ~.,
                          data = heartdisease,
                          cost.outcomes = heart.cost.outcomes,
                          cost.cues = heart.cue.cost,
                          goal = "bacc",
                          goal.chase = "bacc")
```

Let's look at the performance of the best performing tree. The best tree has `bacc` = 0.812 and `cost` = 252

```{r}
summary(heart.costA.fft)$train[1,]
```

Now, we can build an FFT that tries to respect these costs:

```{r}
heart.costB.fft <- FFTrees(formula = diagnosis ~.,
                          data = heartdisease,
                          cost.outcomes = heart.cost.outcomes,
                          cost.cues = heart.cue.cost,
                          goal = "cost",
                          goal.chase = "cost")
```


Here's it's performance: it has a slightly lower balanced accuracy of `bacc = 0.732`, but a much loser cost of `cost = 154`

```{r}
summary(heart.costB.fft)$train[1,]
```



## `"max"` `"zigzag"`

The max and zigzag algorithms are described in @martignon2008categorization.




### References
