---
title: "Explore_NAs_basic"
author: "Jelena Meyer"
date: "2023-06-12"
output: html_document
---

<<<<<<< HEAD
=======

```{r setup, eval = FALSE}
knitr::opts_chunk$set(echo = FALSE)

# load packages
# library(FFTrees)

# load replace_values function
replace_values_copy <- function(data, cols, amount, replacement = NA, levels_amount = NULL) {


  # Verify inputs ----
  
  testthat::expect_true(is.data.frame(data), info = "Data should be a dataframe.") # check that data is a dataframe.
  testthat::expect_true(all(cols %in% names(data)), info = "All column names should be present in the data.") # check that columns is/are all variables in the dataframe.
  testthat::expect_true(length(cols) == length(amount), info = "Number of columns and percentages have to match.") # check if columns and percentages are of the same length.
  testthat::expect_true(all(is.numeric(amount) & (amount >= 0) ), info = "All amounts should be numeric values.") # check that percentages are all numbers between 0 and 1, or higher (than treat as number of to be replaced values.
  testthat::expect_true(is.character(replacement) | is.numeric(replacement) | is.logical(replacement) | is.na(replacement), info = "Replacement value should be of a valid data type (character, nummeric, logical or NA).") # check that replacement is a valid data type.
  testthat::expect_true(is.null(levels_amount) | is.list(levels_amount), info = "levels_amount should be a list") # check that levels_amount is a list if it's not NULL.
  if (!is.null(levels_amount)) {
    testthat::expect_true(all(names(levels_amount) %in% cols), info = "All names in levels_amount have to correspond to a column in the data.") # check if all elements in levels_amount correspond to a column in the data frame.
    for (col in names(levels_amount)) {
      if (is.factor(data[[col]])) {
        testthat::expect_true(all(names(levels_amount[[col]]) %in% levels(data[[col]])), info = "All levels in levels_amount have to be present in the corresponding column of the data.") # check if all sub-elements in each element are valid levels in the corresponding factor variable.

         }

      else if (is.character(data[[col]])) {
        testthat::expect_true(all(names(levels_amount[[col]]) %in% unique(data[[col]])), info = "All levels in levels_amount have to be present in the corresponding column of the data.")

         }

     }

  }


  # loop function over all columns that are inserted in form of a vector and corresponding percentages:
  for (i in seq_along(cols)) {

    col <- cols[i]
    perc <- amount[i]


    # Check if the specific column is a factor and has levels_amount defined:
    if (is.factor(data[[col]]) || is.character(data[[col]]) && !is.null(levels_amount) && col %in% names(levels_amount)) {

      # if all these conditions apply for the column, code is executed:
      # get list of different replacement percentages of levels in column:
      lev_amount <- levels_amount[[col]]

      # loop over levels in current column:
      for (cat in names(lev_amount)) {

        # get replacement percentage for current category:
        replace_perc <- lev_amount[[cat]]

        # get rows in column for current category:
        rows <- which(data[[col]] == cat)

        # Calculate how many values should be replaced in category:
        num_replace <-  ifelse(amount < 1, round(replace_perc * length(rows), 0), amount)


        # Use sample to replace specified percentage of category with replacement input:
        replace_rows <- sample(rows, size = num_replace [1],  replace = FALSE)
        data[replace_rows, col] <- replacement

      }

    } else {

      # Calculate how many values should be replaced:
      num_values <- nrow(data)
      num_replace <-  ifelse(amount < 1, round(amount * num_values, 0), amount)

      # Use sample to replace specified percentage with replacement input:
      replace_rows <- sample(1:num_values,  size = num_replace [1],  replace = FALSE)
      data[replace_rows, col] <- replacement

    }

  }

  # Output: ----

  return(data) # as data frame.

} # replace_values().

```


>>>>>>> 7f858ad (fixing bugs)
```{r Baseline}

#Build tree for normal data set:
heart.fft <- FFTrees(formula = diagnosis ~.,
                     data = heart.train,
                     main = "Heart Disease",
                     decision.labels = c("Healthy", "Diseased"))

```

```{r Data-with-25pc-NAs}

# Preparations:

# get all names of variables but exclude criterion: 
column_names_heart <- colnames(heart.train)
column_names_heart <- column_names_heart[column_names_heart != 'diagnosis']


# in total 25%/50%/75% should be NAs in the end, split equally over all variables: 
num_columns <- length(column_names_heart)
pc_per_row_25 <- 0.25

# create vector with that pc for number of rows:
pc_vector_25 <- rep(pc_per_row_25, times = num_columns)

# apply replace_values function and thereby create new data frame:

# set seed for being able to replicate results:
set.seed(13)

# 25% NAs in data set:
heart_NA_25 <- replace_values(data = heart.train, cols = column_names_heart, amount = pc_vector_25)

<<<<<<< HEAD

=======
>>>>>>> 7f858ad (fixing bugs)
# build best tree. 25% NAs (standard function)
heart.fft_NA_25 <- FFTrees(formula = diagnosis ~.,
                           data = heart_NA_25,
                           main = "Heart Disease",
                           decision.labels = c("Healthy", "Diseased"))


# Report on NA cases: 
tree_level_outcome_vars <- c(1:3, 7:12, 25:30)
heart.fft_NA_25$trees$level_stats$train[ , tree_level_outcome_vars]
  
plot(heart.fft_NA_25)
```

<<<<<<< HEAD

=======
>>>>>>> 7f858ad (fixing bugs)
```{r Data-with-50pc-NAs}

# Preparations:

# get all names of variables but exclude criterion: 
column_names_heart <- colnames(heart.train)
column_names_heart <- column_names_heart[column_names_heart != 'diagnosis']

# in total 50% should be NAs in the end, split equally over all variables: 
num_columns <- length(column_names_heart)
<<<<<<< HEAD
pc_per_row_50 <- 0.50
=======
pc_per_row_50 <- 0.50 
>>>>>>> 7f858ad (fixing bugs)

# create vector with that pc for number of rows:
pc_vector_50 <- rep(pc_per_row_50, times = num_columns)

# apply replace_values function and thereby create new data frame:

# set seed for being able to replicate results:
set.seed(1313)

# 50% NAs in data set:
heart_NA_50 <- replace_values(data = heart.train, cols = column_names_heart, amount = pc_vector_50)



# build best tree, 50% NAs (standard function)
heart.fft_NA_50 <- FFTrees(formula = diagnosis ~.,
                     data = heart_NA_50,
                     main = "Heart Disease",
                     decision.labels = c("Healthy", "Diseased"),
                     do.comp = FALSE)


# Report on NA cases: 
tree_level_outcome_vars <- c(1:3, 7:12, 25:30)
heart.fft_NA_50$trees$level_stats$train[ , tree_level_outcome_vars]
  

plot(heart.fft_NA_50)



```


<<<<<<< HEAD
```{r Data-with-75pc-NAs}
=======
```{r  Data-with-75pc-NAs}
>>>>>>> 7f858ad (fixing bugs)

# Preparations:

# get all names of variables but exclude criterion: 
column_names_heart <- colnames(heart.train)
column_names_heart <- column_names_heart[column_names_heart != 'diagnosis']

# in total 75% should be NAs in the end, split equally over all variables: 
num_columns <- length(column_names_heart)
pc_per_row_75 <- 0.75

# create vector with that pc for number of rows:
pc_vector_75 <- rep(pc_per_row_75, times = num_columns)

# apply replace_values function and thereby create new data frame:

# set seed for being able to replicate results:
set.seed(1333)

# 75% NAs in data set:
heart_NA_75 <- replace_values(data = heart.train, cols = column_names_heart, amount = pc_vector_75)



# build best tree, 75% NAs (standard function)
heart.fft_NA_75 <- FFTrees(formula = diagnosis ~.,
                     data = heart_NA_75,
                     main = "Heart Disease",
                     decision.labels = c("Healthy", "Diseased"),
                     do.comp = FALSE)



# Report on NA cases: 
tree_level_outcome_vars <- c(1:3, 7:12, 25:30)
heart.fft_NA_75$trees$level_stats$train[ , tree_level_outcome_vars]
  
plot(heart.fft_NA_75)


```


<<<<<<< HEAD
```{r Data-with-100pc-NAs-in-TESTDATA}
=======
```{r  Data-with-100pc-NAs}
>>>>>>> 7f858ad (fixing bugs)

# Preparations:

# get all names of variables but exclude criterion: 
column_names_heart <- colnames(heart.test)
column_names_heart <- column_names_heart[column_names_heart != 'diagnosis']

# in total 100% per row should be NAs in the end: 
per_row_100 <- nrow(heart.test)


# create vector with number of rows:
num_columns <- length(column_names_heart)
vector_100 <- rep(per_row_100, times = num_columns)

# apply replace_values function and thereby create new data frames:

# set seed for being able to replicate results:
set.seed(13333)

# 100% NAs in data set:
<<<<<<< HEAD
heart_NA_100 <- replace_values(data = heart.test, cols = column_names_heart, amount = vector_100)
=======
heart_NA_100 <- replace_values(data = heart.train, cols = column_names_heart, amount = vector_100)
>>>>>>> 7f858ad (fixing bugs)

# Check: 
sum(is.na(heart_NA_100)) == (ncol(heart_NA_100) - 1) * nrow(heart_NA_100)

# build best tree, 100% NAs (standard function)
heart.fft_NA_100 <- FFTrees(formula = diagnosis ~.,
<<<<<<< HEAD
                     data = heart.train,
                     data.test = heart_NA_100,
                     main = "Heart Disease",
                     decision.labels = c("Healthy", "Diseased"),
                     do.comp = FALSE)
=======
                    data = heart_NA_100,
                    main = "Heart Disease",
                    decision.labels = c("Healthy", "Diseased"),
                    quiet = list(ini = FALSE, fin = FALSE, mis = FALSE, set = TRUE), 
                    do.comp = FALSE)
>>>>>>> 7f858ad (fixing bugs)



# Report on NA cases: 
tree_level_outcome_vars <- c(1:3, 7:12, 25:30)
heart.fft_NA_100$trees$level_stats$test[ , tree_level_outcome_vars]
  
plot(heart.fft_NA_100, data = "test") # it shows the expected behaviour.


```

