---
title: "NA_5_10_pc_NA"
author: "JM and HN"
date: "2023-06-21"
output: html_document
---

# Define Loop functions (for multiple simulations) ----

"_T" because here, the functions are defined differently: A tree is always built with the new data set. New combinations are therefore possible, (e.g. NA can be used as category in character variables). We thereby allow for compensation and hence have slightly better outcomes than when applying the original tree to data with a certain percentage of missings. 

```{r define-loop-NA-T-function }

loop_NA_T <- function(data, data_name, criterion, pc, x, fft_df, x_stats_1){
  
  
  # Prepare for loop:
  # Parameters:
  n_per_NA <- 10
  n_sims <- n_per_NA
  
  # initialize data structure to collect results:
  results <- data.frame(matrix(NA, nrow = 1 + n_sims, ncol = 26))
  
  # results:
  n_col <- ncol(x_stats_1)
  results[1, 1:n_col] <- x_stats_1
  
  
  # Prepare vectors as input for replace_values function:  
  # get all names of variables but exclude criterion: 
  col_names <- names(data)
  col_names <- col_names[col_names != criterion ]
  
  # build vector with NA percentage: 
  num_cols <- length(col_names)
  pc_per_row <- pc
  
  # create vector with that pc for number of rows:
  pc_vec <- rep(pc_per_row, times = num_cols)
  
  # set seed for reproducability:
  set.seed(13)
  
  # Loop 1: ---- 
  for (i in 1:n_sims){
    
    # Add NAs to data:
    data_NA <- replace_values(data = data, 
                              cols = col_names, 
                              amount = pc_vec, 
                              replacement = NA, 
                              levels_amount = NULL)
    
    
    # Apply tree definition to (OLD/NBW) data:
    y <- FFTrees(formula = as.formula(paste(criterion,"~.")),
                 data = data_NA,
                 main = data_name,
                 quiet = TRUE, 
                 do.comp = FALSE)
    
    # Collect results:
    
    # a) tree performance: 
    y_summary <- summary(y)
    y_stats <- y_summary$stats$train[1, ]
    
    # b) level stats:
    y_level_stats <- y$trees$level_stats$train[1,]
    NA_stats <- y_level_stats[nrow(y_level_stats) , c(25 :30)]
    
    # Combine:
    all_data_df <- cbind(y_stats, NA_stats)
    # dim(all_data_df)
    
    
    # Collect results:
    results[(1 + i), ] <- all_data_df
    
  } # for loop. 
  
  
  # After loop:
  
  # Add column names to data frame:
  names(results) <- names(all_data_df)
  
  
  # Output: ----
  return(results) # as data frame.
  
} # loop_NA().

```


```{r define-loop-pc-T-function}

loop_pc_T <- function(data, data_name, criterion, several_pcs, x, fft_df, x_stats_1){
  
  # Initialize list to store results:
  results_dataframes <-  list()
  
  # length of vector, to know when to stop:
  n <- length(several_pcs)
  
  # Loop 2:
  for (j in 1:n) {
    
    # use function loop_NA and give name to each data frame:
    result_name <- paste0(data_name, "_", several_pcs[j]*100, "pc")
    results_dataframes[[result_name]] <- loop_NA_T(data, data_name, criterion, several_pcs[j], x, fft_df, x_stats_1)
    
  } # for loop.
  
  # Output: ----
  return(results_dataframes)# as list.
  
} # loop_pc.

```


```{r define-loop-datasets-function}

loop_datasets_T <- function(several_data, data_names, several_criterion, several_pcs){
  
  # Preparations:
  # length of vector, to know when to stop:
  n <- length(several_data)
  
  
  # Initialize list to store results:
  results_list_dataframes <-  list()
  
  # Initialize list to store results of Baseline-pc:
  baseline_pc <- numeric(n)
  
  
  # Initialize two numeric vectors to store results of Baseline-pc:
  bl <- numeric(n)
  pred_benchmark <- numeric(n)
  
  # Loop 3:
  for (k in 1:n) {
    
    # build tree with original data for each dataset:
    # get name of data set:
    name <- data_names[k]
    
    #Build tree for normal data set:
    x <- FFTrees(formula = as.formula(paste(several_criterion[k],"~.")),
                 data = several_data[[k]],
                 main = name,
                 goal = "bacc",
                 goal.chase = "bacc",
                 goal.threshold = "bacc")
    
    # Performance:
    x_summary <- summary(x)
    x_stats_1 <- x_summary$stats$train[1, ]
    
    # get tree definitions;
    x_fft_df <- get_fft_df(x)
    fft_df <- x_fft_df[1, ]  # best training tree
    
  
    # use function loop_pc:
    results_list_dataframes[[data_names[k]]] <-loop_pc_T(several_data[[k]], data_names[k], several_criterion[k], several_pcs, x, fft_df, x_stats_1)
    
    
    # describe data for further table:
    dd <- describe_data(several_data[[k]], 
                        data_name = name, 
                        criterion_name = several_criterion[k], 
                        baseline_value = TRUE)
    
    # Baseline:
    bl[k] <- dd$Baseline_pct
    
    # Minimum prediction performance (i.e., majority of criterion):
    pred_benchmark[k] <- max(bl[k], 100 - bl[k])/100
    
    # store baseline_pc:
    baseline_pc[k] <- pred_benchmark[k]
    
    
  } # for loop.
  
  # Output: ----
  return(list(results = results_list_dataframes, baseline = baseline_pc)) # as list.
  
} # loop_datasets.

```

# Apply

```{r apply}
# original data set, compare to data sets who built trees individually (with 5, 10% NAs):

try <- loop_datasets_T(several_data = list(heartdisease),
                       data_names =  c("heartdisease"),
                       several_criterion = c("diagnosis"),
                       several_pcs = c(0.05, 0.10)
                       )

# heartdisease:
heart_NA_05 <- try$results$heartdisease[["heartdisease_5pc"]]
heart_NA_10 <- try$results$heartdisease[["heartdisease_10pc"]]
```


```{r key results}

# Collect key results (in df): 
  
key_results <- data.frame(matrix(NA, nrow = 3, ncol = 4))
names(key_results) <- c("pc_NA", "mn_acc", "mn_bacc", "mn_mcu")

# Baseline performance:
key_results[1, "pc_NA"]   <- 0
key_results[1, "mn_acc"]  <- heart_NA_05$acc[1]
key_results[1, "mn_bacc"] <- heart_NA_05$bacc[1]
key_results[1, "mn_mcu"]  <- heart_NA_05$mcu[1]

# Row 2: 5% NA values:
key_results[2, "pc_NA"]   <- 5
key_results[2, "mn_acc"]  <- mean(heart_NA_05$acc[-1])
key_results[2, "mn_bacc"] <- mean(heart_NA_05$bacc[-1])
key_results[2, "mn_mcu"]  <- mean(heart_NA_05$mcu[-1])


# Row 3: 10% NA values:
key_results[3, "pc_NA"]   <- 10
key_results[3, "mn_acc"]  <- mean(heart_NA_10$acc[-1])
key_results[3, "mn_bacc"] <- mean(heart_NA_10$bacc[-1])
key_results[3, "mn_mcu"]  <- mean(heart_NA_10$mcu[-1])

key_results
```


```{r plot results}

# Upper benchmark: Best FFT without missing data: 
plot(x = key_results$pc_NA, 
     y = rep(key_results$mn_acc[1], nrow(key_results)), 
     type = "l", 
     col = unikn::Seegruen, lwd = 1, lty = 2, 
     ylim = c(0, 1),
     xlab = "Percent NA values", 
     ylab = "Mean accuracy"
)
grid()

# Lower benchmark pred_benchmark (i.e., majority of criterion):
lines(x = key_results$pc_NA, 
      y = rep(try$baseline, nrow(key_results)), 
      type = "l", 
      col = unikn::Bordeaux, lwd = 1, lty = 2 
      # ylim = c(0, 1),
      # xlab = "Percent NA values", 
      # ylab = "Mean accuracy"
)

# Actual accuracy:
lines(x = key_results$pc_NA, 
      y = key_results$mn_acc, 
      type = "b", 
      col = unikn::Seeblau, lwd = 2
      # ylim = c(0, 1),
      # xlab = "Percent NA values", 
      # ylab = "Mean accuracy"
)

title(main = "Mean accuracy given missing values - heartdisease data")
```


