---
title: "NA_explore_nocompensation"
author: "Jelena Meyer"
date: "2023-06-14"
output: html_document
---

# Additional interest: No compensation
Besides investigating the performance of newly built trees, we are also interested in the trees performance using the tree that was the best when there was no missing data in the data set yet.

```{r Baseline}

#Build tree for normal data set:
heart.fft <- FFTrees(formula = diagnosis ~.,
                     data = heartdisease,
                     main = "Heart Disease",
                     decision.labels = c("Healthy", "Diseased"))


# get definition of best train.tree:
def_best <- inwords(heart.fft, tree = 1)

```


```{r Tree-performance-not-allowing-for-compensation-5pc-NAs}


# Preparations:

# get all names of variables but exclude criterion: 
column_names_heart <- colnames(heartdisease)
column_names_heart <- column_names_heart[column_names_heart != 'diagnosis']


# in total 5%/50%/75% should be NAs in the end, split equally over all variables: 
num_columns <- length(column_names_heart)
pc_per_row_5 <- 0.05

# create vector with that pc for number of rows:
pc_vector_5 <- rep(pc_per_row_5, times = num_columns)

# apply replace_values function and thereby create new data frame:

# set seed for being able to replicate results:
set.seed(13)

# 5% NAs in data set:
heart_NA_5 <- replace_values(data = heartdisease, cols = column_names_heart, amount = pc_vector_5)

# Define trees with new data but old best tree:

# data NA 5%
heart.fft_NA_5_oldbest <- FFTrees(formula = diagnosis ~.,
                                   data = heart_NA_5,
                                   my.tree = def_best,
                                   main = "Heart Disease",
                                   decision.labels = c("Healthy", "Diseased"))

plot(heart.fft_NA_5_oldbest) # with compensation acc = 81%, without worse!


```



```{r Tree-performance-not-allowing-for-compensation-10pc-NAs}


# Preparations:

# get all names of variables but exclude criterion: 
column_names_heart <- colnames(heartdisease)
column_names_heart <- column_names_heart[column_names_heart != 'diagnosis']


# in total 10%/50%/75% should be NAs in the end, split equally over all variables: 
num_columns <- length(column_names_heart)
pc_per_row_10 <- 0.10

# create vector with that pc for number of rows:
pc_vector_10 <- rep(pc_per_row_10, times = num_columns)

# apply replace_values function and thereby create new data frame:

# set seed for being able to replicate results:
set.seed(13)

# 10% NAs in data set:
heart_NA_10 <- replace_values(data = heartdisease, cols = column_names_heart, amount = pc_vector_10)

# Define trees with new data but old best tree:

# data NA 10%
heart.fft_NA_10_oldbest <- FFTrees(formula = diagnosis ~.,
                                   data = heart_NA_10,
                                   my.tree = def_best,
                                   main = "Heart Disease",
                                   decision.labels = c("Healthy", "Diseased"))

plot(heart.fft_NA_10_oldbest) # with compensation acc = 81%, without worse!


```


```{r Tree-performance-not-allowing-for-compensation-25pc-NAs}


# Preparations:

# get all names of variables but exclude criterion: 
column_names_heart <- colnames(heartdisease)
column_names_heart <- column_names_heart[column_names_heart != 'diagnosis']


# in total 25%/50%/75% should be NAs in the end, split equally over all variables: 
num_columns <- length(column_names_heart)
pc_per_row_25 <- 0.25

# create vector with that pc for number of rows:
pc_vector_25 <- rep(pc_per_row_25, times = num_columns)

# apply replace_values function and thereby create new data frame:

# set seed for being able to replicate results:
set.seed(13)

# 25% NAs in data set:
heart_NA_25 <- replace_values(data = heartdisease, cols = column_names_heart, amount = pc_vector_25)

# Define trees with new data but old best tree:

# data NA 25%
heart.fft_NA_25_oldbest <- FFTrees(formula = diagnosis ~.,
                                   data = heart_NA_25,
                                   my.tree = def_best,
                                   main = "Heart Disease",
                                   decision.labels = c("Healthy", "Diseased"))

plot(heart.fft_NA_25_oldbest) # with compensation acc = 81%, without worse!


```


```{r Tree-performance-not-allowing-for-compensation-50pc-NAs}

# Preparations:

# get all names of variables but exclude criterion: 
column_names_heart <- colnames(heartdisease)
column_names_heart <- column_names_heart[column_names_heart != 'diagnosis']

# in total 50% should be NAs in the end, split equally over all variables: 
num_columns <- length(column_names_heart)
pc_per_row_50 <- 0.50

# create vector with that pc for number of rows:
pc_vector_50 <- rep(pc_per_row_50, times = num_columns)

# apply replace_values function and thereby create new data frame:

# set seed for being able to replicate results:
set.seed(133)

# 50% NAs in data set:
heart_NA_50 <- replace_values(data = heartdisease, cols = column_names_heart, amount = pc_vector_50)



# build best tree, 50% NAs (standard function)
heart.fft_NA_50_oldbest <- FFTrees(formula = diagnosis ~.,
                                   data = heart_NA_50,
                                   my.tree = def_best,
                                   main = "Heart Disease",
                                   decision.labels = c("Healthy", "Diseased"),
                                   do.comp = FALSE)

plot(heart.fft_NA_50_oldbest) # with compensation acc = 76%, without worse!


```


```{r Tree-performance-not-allowing-for-compensation-75pc-NAs}

# Preparations:

# get all names of variables but exclude criterion: 
column_names_heart <- colnames(heartdisease)
column_names_heart <- column_names_heart[column_names_heart != 'diagnosis']

# in total 75% should be NAs in the end, split equally over all variables: 
num_columns <- length(column_names_heart)
pc_per_row_75 <- 0.75

# create vector with that pc for number of rows:
pc_vector_75 <- rep(pc_per_row_75, times = num_columns)

# apply replace_values function and thereby create new data frame:

# set seed for being able to replicate results:
set.seed(1333)

# 75% NAs in data set:
heart_NA_75 <- replace_values(data = heartdisease, cols = column_names_heart, amount = pc_vector_75)


# build best tree, 75% NAs (standard function)
heart.fft_NA_75_oldbest <- FFTrees(formula = diagnosis ~.,
                                   data = heart_NA_75,
                                   my.tree = def_best,
                                   main = "Heart Disease",
                                   decision.labels = c("Healthy", "Diseased"),
                                   do.comp = FALSE)


plot(heart.fft_NA_75_oldbest)  # with compensation acc = 66%, without worse!

```



```{r Tree-performance-not-allowing-for-compensation-75pc-NAs}

# Preparations:

# get all names of variables but exclude criterion: 
column_names_heart <- colnames(heartdisease)
column_names_heart <- column_names_heart[column_names_heart != 'diagnosis']

# in total 75% should be NAs in the end, split equally over all variables: 
num_columns <- length(column_names_heart)
pc_per_row_100 <- 1

# create vector with that pc for number of rows:
pc_vector_100 <- rep(pc_per_row_100, times = num_columns)

# apply replace_values function and thereby create new data frame:

# set seed for being able to replicate results:
set.seed(1333)

# 75% NAs in data set:
heart_NA_100 <- replace_values(data = heartdisease, cols = column_names_heart, amount = pc_vector_100)


# build best tree, 100% NAs (standard function)
heart.fft_NA_100_oldbest <- FFTrees(formula = diagnosis ~.,
                                   data = heart_NA_100,
                                   my.tree = def_best,
                                   main = "Heart Disease",
                                   decision.labels = c("Healthy", "Diseased"),
                                   do.comp = FALSE)


plot(heart.fft_NA_100_oldbest)  # with compensation acc = 66%, without worse!

```




In all cases the performance of the trees was worse when not allowing for missing data versus when compensating was allowed.
Probably dependent on how well different variables predict outcome variable.
