---
title: "NA_General"
author: "JM and HN"
date: "2023-06-21"
output: html_document
---

# Define Loop functions (for multiple simulations) ----


```{r define-loop-function}

loop_NA <- function(data, criterion, pc, x, fft_df, x_stats_1){


# Prepare for loop:
  # Parameters:
  n_per_NA <- 10
  n_sims <- n_per_NA

# initialize data structure to collect results:
results <- data.frame(matrix(NA, nrow = 1 + n_sims, ncol = 26))

# results:
n_col <- ncol(x_stats_1)
results[1, 1:n_col] <- x_stats_1

  
# Prepare vectors as input for replace_values function:  
   # get all names of variables but exclude criterion: 
   col_names <- names(data)
   col_names <- col_names[col_names != criterion ]
  
   # build vector with NA percentage: 
   num_cols <- length(col_names)
   pc_per_row <- pc

   # create vector with that pc for number of rows:
   pc_vec <- rep(pc_per_row, times = num_cols)

   # set seed for reproducability
   set.seed(13)
  
# Loop 1: ---- 
for (i in 1:n_sims){

# Add NAs to data:
data_NA <- replace_values(data = data, 
                          cols = col_names, 
                          amount = pc_vec, 
                          replacement = NA, 
                          levels_amount = NULL)

  
# Apply tree definition to (OLD/NBW) data:
y <- FFTrees(object = x,
             tree.definitions = fft_df,
             data = data_NA, 
             quiet = TRUE, 
             do.comp = FALSE)

# Collect results:

# a) tree performance: 
y_summary <- summary(y)
y_stats <- y_summary$stats$train

# b) level stats:
y_level_stats <- y$trees$level_stats$train
NA_stats <- y_level_stats[nrow(y_level_stats) , c(25 :30)]

# Combine:
all_data_df <- cbind(y_stats, NA_stats)
# dim(all_data_df)


# Collect results:
results[(1 + i), ] <- all_data_df

} # for loop. 

  
# After loop:

# Add column names to data frame:
names(results) <- names(all_data_df)

       
# Output: ----
return(results) # as data frame.

} # loop_NA().
```

```{r define-loop-pc-function}

# as input a vector with different pcs for which you want data:

loop_pc <- function(data, data_name, criterion, several_pcs, x, fft_df, x_stats_1){

  # Initialize list to store results
  results_dataframes <-  list()

  # length of vector, to know when to stop
  n <- length(several_pcs)

  # Loop 2:
  for (j in 1:n) {

    # use function loop_NA
    result_name <- paste0(data_name, "_", several_pcs[j]*100, "pc")
    results_dataframes[[result_name]] <-loop_NA(data, criterion, several_pcs[j], x, fft_df, x_stats_1)

  } # for loop.

  # Output: ----
  return(results_dataframes)# as list.

} # loop_pc.

```

```{r define-loop-datasets-function}

# as input a vector with different data sets for which you want data:

loop_datasets <- function(several_data, data_names, several_criterion, several_pcs){

  # Preparations:
  # length of vector, to know when to stop
  n <- length(several_data)


  # Initialize list to store results
  results_list_dataframes <-  list()
  
  # Initialize list to store results of Baseline-pc:
  baseline_pc <- numeric(n)

  
  # Initialize two nummeric vectors to store results of Baseline-pc:
  bl <- numeric(n)
  pred_benchmark <- numeric(n)
  
  # Loop 3:
  for (k in 1:n) {

    # build tree with original data for each dataset:
    # get name of data set:
    name <- data_names[k]

    #Build tree for normal data set:
    x <- FFTrees(formula = as.formula(paste(several_criterion[k],"~.")),
                 data = several_data[[k]],
                 main = name,
                 goal = "bacc",
                 goal.chase = "bacc",
                 goal.threshold = "bacc")

    # Performance:
    x_summary <- summary(x)
    x_stats_1 <- x_summary$stats$train[1, ]

    # get tree definitions;
    x_fft_df <- get_fft_df(x)
    fft_df <- x_fft_df[1, ]  # best training tree



    # use function loop_pc
    results_list_dataframes[[data_names[k]]] <-loop_pc(several_data[[k]], data_names[k], several_criterion[k], several_pcs, x, fft_df, x_stats_1)

    
    
    # describe data for further table:
    dd <- describe_data(several_data[[k]], 
                    data_name = name, 
                    criterion_name = several_criterion[k], 
                    baseline_value = TRUE)

    # Baseline:
    bl[k] <- dd$Baseline_pct

    # Minimum prediction performance (i.e., majority of criterion):
    pred_benchmark[k] <- max(bl[k], 100 - bl[k])/100
    
    # store baseline_pc:
    baseline_pc[k] <- pred_benchmark[k]

    
  } # for loop.

  # Output: ----
  return(list(results = results_list_dataframes, baseline = baseline_pc)) # as list.

} # loop_datasets.

```





# Run simulations, make graph: ----

```{r run-simulation}

dfs <- loop_datasets(several_data = list(heartdisease, titanic, breastcancer),
                      data_names = c("heartdisease", "titanic", "breastcancer"), # expl: the get_object_name <- function(x){deparse(substitute(x)) does not work this way, so we have to manually define names.
                      several_criterion = c("diagnosis", "survived", "diagnosis"),
                      several_pcs = c(0.25, 0.50, 0.75, 1))


# assign names to output dfs:
# heartdisease:
heart_NA_25 <- dfs$results$heartdisease[["heartdisease_25pc"]]
heart_NA_50 <- dfs$results$heartdisease[["heartdisease_50pc"]]
heart_NA_75 <- dfs$results$heartdisease[["heartdisease_75pc"]]
heart_NA_100 <- dfs$results$heartdisease[["heartdisease_100pc"]]


# titanic:
titanic_NA_25 <- dfs$results$titanic[["titanic_25pc"]]
titanic_NA_50 <- dfs$results$titanic[["titanic_50pc"]]
titanic_NA_75 <- dfs$results$titanic[["titanic_75pc"]]
titanic_NA_100 <- dfs$results$titanic[["titanic_100pc"]]


# breastcancer:
breastcancer_NA_25 <- dfs$results$breastcancer[["breastcancer_25pc"]]
breastcancer_NA_50 <- dfs$results$breastcancer[["breastcancer_50pc"]]
breastcancer_NA_75 <- dfs$results$breastcancer[["breastcancer_75pc"]]
breastcancer_NA_100 <- dfs$results$breastcancer[["breastcancer_100pc"]]


# dfs$baseline => 0.5412541 0.6769650 0.6500732
 

```

# for heartdisease: ----

Collect key results (in df): 

```{r key-results}
key_results <- data.frame(matrix(NA, nrow = 5, ncol = 4))
names(key_results) <- c("pc_NA", "mn_acc", "mn_bacc", "mn_mcu")

# Baseline performance:
key_results[1, "pc_NA"]   <- 0
key_results[1, "mn_acc"]  <- heart_NA_25$acc[1]
key_results[1, "mn_bacc"] <- heart_NA_25$bacc[1]
key_results[1, "mn_mcu"]  <- heart_NA_25$mcu[1]

# Row 2: 25% NA values:
key_results[2, "pc_NA"]   <- 25
key_results[2, "mn_acc"]  <- mean(heart_NA_25$acc[-1])
key_results[2, "mn_bacc"] <- mean(heart_NA_25$bacc[-1])
key_results[2, "mn_mcu"]  <- mean(heart_NA_25$mcu[-1])


# Row 3: 50% NA values:
key_results[3, "pc_NA"]   <- 50
key_results[3, "mn_acc"]  <- mean(heart_NA_50$acc[-1])
key_results[3, "mn_bacc"] <- mean(heart_NA_50$bacc[-1])
key_results[3, "mn_mcu"]  <- mean(heart_NA_50$mcu[-1])

# Row 4: 75% NA values:
key_results[4, "pc_NA"]   <- 75
key_results[4, "mn_acc"]  <- mean(heart_NA_75$acc[-1])
key_results[4, "mn_bacc"] <- mean(heart_NA_75$bacc[-1])
key_results[4, "mn_mcu"]  <- mean(heart_NA_75$mcu[-1])

# Basically a sanity check: 
# Row 5: 100% NA values:
key_results[5, "pc_NA"]   <- 100
key_results[5, "mn_acc"]  <- mean(heart_NA_100$acc[-1])
key_results[5, "mn_bacc"] <- mean(heart_NA_100$bacc[-1])
key_results[5, "mn_mcu"]  <- mean(heart_NA_100$mcu[-1])


key_results
```

Plot results:

```{r plot-key-results}
# Upper benchmark: Best FFT without missing data: 
plot(x = key_results$pc_NA, 
     y = rep(key_results$mn_acc[1], nrow(key_results)), 
     type = "l", 
     col = unikn::Seegruen, lwd = 1, lty = 2, 
     ylim = c(0, 1),
     xlab = "Percent NA values", 
     ylab = "Mean accuracy"
     )
grid()
# Lower benchmark pred_benchmark (i.e., majority of criterion):
lines(x = key_results$pc_NA, 
      y = rep(dfs$baseline[1], nrow(key_results)), 
      type = "l", 
      col = unikn::Bordeaux, lwd = 1, lty = 2 
      # ylim = c(0, 1),
      # xlab = "Percent NA values", 
      # ylab = "Mean accuracy"
)
# Actual accuracy:
lines(x = key_results$pc_NA, 
      y = key_results$mn_acc, 
      type = "b", 
      col = unikn::Seeblau, lwd = 2
      # ylim = c(0, 1),
      # xlab = "Percent NA values", 
      # ylab = "Mean accuracy"
)
title(main = "Mean accuracy given missing values - heartdisease data")
```


# for titanic: ----

Collect key results (in df): 

```{r key-results}
key_results <- data.frame(matrix(NA, nrow = 5, ncol = 4))
names(key_results) <- c("pc_NA", "mn_acc", "mn_bacc", "mn_mcu")

# Baseline performance:
key_results[1, "pc_NA"]   <- 0
key_results[1, "mn_acc"]  <- titanic_NA_25$acc[1]
key_results[1, "mn_bacc"] <- titanic_NA_25$bacc[1]
key_results[1, "mn_mcu"]  <- titanic_NA_25$mcu[1]

# Row 2: 25% NA values:
key_results[2, "pc_NA"]   <- 25
key_results[2, "mn_acc"]  <- mean(titanic_NA_25$acc[-1])
key_results[2, "mn_bacc"] <- mean(titanic_NA_25$bacc[-1])
key_results[2, "mn_mcu"]  <- mean(titanic_NA_25$mcu[-1])


# Row 3: 50% NA values:
key_results[3, "pc_NA"]   <- 50
key_results[3, "mn_acc"]  <- mean(titanic_NA_50$acc[-1])
key_results[3, "mn_bacc"] <- mean(titanic_NA_50$bacc[-1])
key_results[3, "mn_mcu"]  <- mean(titanic_NA_50$mcu[-1])

# Row 4: 75% NA values:
key_results[4, "pc_NA"]   <- 75
key_results[4, "mn_acc"]  <- mean(titanic_NA_75$acc[-1])
key_results[4, "mn_bacc"] <- mean(titanic_NA_75$bacc[-1])
key_results[4, "mn_mcu"]  <- mean(titanic_NA_75$mcu[-1])

# Basically a sanity check: 
# Row 5: 100% NA values:
key_results[5, "pc_NA"]   <- 100
key_results[5, "mn_acc"]  <- mean(titanic_NA_100$acc[-1])
key_results[5, "mn_bacc"] <- mean(titanic_NA_100$bacc[-1])
key_results[5, "mn_mcu"]  <- mean(titanic_NA_100$mcu[-1])


key_results
```

Plot results:

```{r plot-key-results}
# Upper benchmark: Best FFT without missing data: 
plot(x = key_results$pc_NA, 
     y = rep(key_results$mn_acc[1], nrow(key_results)), 
     type = "l", 
     col = unikn::Seegruen, lwd = 1, lty = 2, 
     ylim = c(0, 1),
     xlab = "Percent NA values", 
     ylab = "Mean accuracy"
     )
grid()
# Lower benchmark pred_benchmark (i.e., majority of criterion):
lines(x = key_results$pc_NA, 
      y = rep(dfs$baseline[2], nrow(key_results)), 
      type = "l", 
      col = unikn::Bordeaux, lwd = 1, lty = 2 
      # ylim = c(0, 1),
      # xlab = "Percent NA values", 
      # ylab = "Mean accuracy"
)
# Actual accuracy:
lines(x = key_results$pc_NA, 
      y = key_results$mn_acc, 
      type = "b", 
      col = unikn::Seeblau, lwd = 2
      # ylim = c(0, 1),
      # xlab = "Percent NA values", 
      # ylab = "Mean accuracy"
)
title(main = "Mean accuracy given missing values - titanic data")
```



# for breastcancer: ----

Collect key results (in df): 

```{r key-results}
key_results <- data.frame(matrix(NA, nrow = 5, ncol = 4))
names(key_results) <- c("pc_NA", "mn_acc", "mn_bacc", "mn_mcu")

# Baseline performance:
key_results[1, "pc_NA"]   <- 0
key_results[1, "mn_acc"]  <- breastcancer_NA_25$acc[1]
key_results[1, "mn_bacc"] <- breastcancer_NA_25$bacc[1]
key_results[1, "mn_mcu"]  <- breastcancer_NA_25$mcu[1]

# Row 2: 25% NA values:
key_results[2, "pc_NA"]   <- 25
key_results[2, "mn_acc"]  <- mean(breastcancer_NA_25$acc[-1])
key_results[2, "mn_bacc"] <- mean(breastcancer_NA_25$bacc[-1])
key_results[2, "mn_mcu"]  <- mean(breastcancer_NA_25$mcu[-1])


# Row 3: 50% NA values:
key_results[3, "pc_NA"]   <- 50
key_results[3, "mn_acc"]  <- mean(breastcancer_NA_50$acc[-1])
key_results[3, "mn_bacc"] <- mean(breastcancer_NA_50$bacc[-1])
key_results[3, "mn_mcu"]  <- mean(breastcancer_NA_50$mcu[-1])

# Row 4: 75% NA values:
key_results[4, "pc_NA"]   <- 75
key_results[4, "mn_acc"]  <- mean(breastcancer_NA_75$acc[-1])
key_results[4, "mn_bacc"] <- mean(breastcancer_NA_75$bacc[-1])
key_results[4, "mn_mcu"]  <- mean(breastcancer_NA_75$mcu[-1])

# Basically a sanity check: 
# Row 5: 100% NA values:
key_results[5, "pc_NA"]   <- 100
key_results[5, "mn_acc"]  <- mean(breastcancer_NA_100$acc[-1])
key_results[5, "mn_bacc"] <- mean(breastcancer_NA_100$bacc[-1])
key_results[5, "mn_mcu"]  <- mean(breastcancer_NA_100$mcu[-1])


key_results
```

Plot results:

```{r plot-key-results}
# Upper benchmark: Best FFT without missing data: 
plot(x = key_results$pc_NA, 
     y = rep(key_results$mn_acc[1], nrow(key_results)), 
     type = "l", 
     col = unikn::Seegruen, lwd = 1, lty = 2, 
     ylim = c(0, 1),
     xlab = "Percent NA values", 
     ylab = "Mean accuracy"
     )
grid()
# Lower benchmark pred_benchmark (i.e., majority of criterion):
lines(x = key_results$pc_NA, 
      y = rep(dfs$baseline[2], nrow(key_results)), 
      type = "l", 
      col = unikn::Bordeaux, lwd = 1, lty = 2 
      # ylim = c(0, 1),
      # xlab = "Percent NA values", 
      # ylab = "Mean accuracy"
)
# Actual accuracy:
lines(x = key_results$pc_NA, 
      y = key_results$mn_acc, 
      type = "b", 
      col = unikn::Seeblau, lwd = 2
      # ylim = c(0, 1),
      # xlab = "Percent NA values", 
      # ylab = "Mean accuracy"
)
title(main = "Mean accuracy given missing values - breastcancer data")
```


# for combination: ---- +++ here now +++ 

```{r}

```





# Issues and potential problems ----


### 1. Using <NA> in categorical predictors

- When creating the basic FFT _without_ any `NA` values, 
the mechanism for _categorical_ predictors (i.e., treating `NA` as a separate `<NA>` category) is not being used or tested.  
Thus, we should create the initial FFT (and corresponding performance baselines) for datasets with _some_ `NA` values (e.g., 10%). DONE!


### 2. Dealing with NA in final predictors

- The current default for coping with `NA` values in final node 
appears to be to assign all cases 
to the _majority_ class (as only 2 types of errors occur): 
`fin_NA_pred = "majority"` (in **FFTrees** v2.0.0). 

- Done: Switched default to `fin_NA_pred = "baseline"` in `fftrees_apply()` (i.e., biased coin flip with both outcomes).


### 3. Value of mcu?

- Why is the final `mcu` value (for simulations with 100% NA values) 
a value of\ 2, rather than\ 3 (for FFT `fft_df` with 3\ nodes)? 

<!-- eof. -->
