---
title: "NA_heartdisease"
author: "jm and hn"
date: "2023-07-05"
output: html_document
---

```{r Preparations}

# get name of data set:
get_object_name <- function(x){

  deparse(substitute(x))

} # get_object_name().

# Check:
# get_object_name(heartdisease)
```

## Data

Describe current data:

```{r describe-data}
dd <- describe_data(heartdisease, 
                    data_name = "heartdisease", 
                    criterion_name = "diagnosis", 
                    baseline_value = TRUE)

# Baseline:
bl <- dd$Baseline_pct
bl

# Minimum prediction performance (i.e., majority of criterion):
pred_benchmark <- max(bl, 100 - bl)/100  # percentage in [0:1]
pred_benchmark
```

Probability of being correct:

```{r randomly-correct-cases}
p <- seq(0, 1, by = .10)  # p of signal/TRUE

p_cum <- (p * p) + ((1 - p) * (1 - p))

plot(x = p, y = p_cum, type = "b", col = unikn::Seegruen)
```



Create 1 FFT for current data (and criteria): 

```{r Get-tree-for-original-data}

# Build tree for normal data set (without NA values):
x <- FFTrees(formula = diagnosis ~.,
             data = heartdisease,
             main = get_object_name(heartdisease),
             goal = "bacc",
             goal.chase = "bacc",
             goal.threshold = "bacc",
             quiet = list(ini = FALSE, fin = FALSE, mis = FALSE, set = FALSE)
)

# Performance:
# print(x)
# plot(x, what = "all")

x_summary <- summary(x)
x_stats_1 <- x_summary$stats$train[1, ]
# x_stats_1

# get tree definitions;
x_fft_df <- get_fft_df(x)
fft_df <- x_fft_df[1, ]  # definition of best training tree
# fft_df
```

Get NA data:

```{r get-NA-data}
# Select dataset:
data <- heartdisease
criterion <- "diagnosis"

pc <- .50

# Prepare vectors as input for replace_values function:  
# get all names of variables but exclude criterion: 
col_names <- names(data)
col_names <- col_names[col_names != criterion ]

# build vector with NA percentage: 
num_cols <- length(col_names)
pc_per_row <- pc

# create vector with that pc for number of rows:
pc_vec <- rep(pc_per_row, times = num_cols)

# Add NAs to data:
data_NA <- replace_values(data = data, 
                          cols = col_names, 
                          amount = pc_vec, 
                          replacement = NA, 
                          levels_amount = NULL)

# data_NA
```

Apply FFT to NA data:

```{r apply-FFT-to-NA-data}
# Apply tree definition to (OLD/NEW NA) data:
y <- FFTrees(object = x,
             tree.definitions = fft_df,
             data = data_NA, 
             quiet = TRUE, 
             do.comp = FALSE)

# FFT:
fft_df

# Data:
cur_cols <- names(data_NA) %in% c("diagnosis", "thal", "cp", "ca")
data_NA[cur_cols]

# Decisions:
y$trees$decisions$train$tree_1
```

Loop function (for multiple simulations): 

```{r define-loop-function}
# Simulation loops:
loop_NA <- function(data, criterion, pc){
  
  # Prepare for loop:
  # Parameters:
  n_per_NA <- 100
  n_sims <- n_per_NA
  
  # initialize data structure to collect results:
  results <- data.frame(matrix(NA, nrow = 1 + n_sims, ncol = 26))
  
  # results:
  n_col <- ncol(x_stats_1)
  results[1, 1:n_col] <- x_stats_1
  
  
  # Prepare vectors as input for replace_values function:  
  # get all names of variables but exclude criterion: 
  col_names <- names(data)
  col_names <- col_names[col_names != criterion ]
  
  # build vector with NA percentage: 
  num_cols <- length(col_names)
  pc_per_row <- pc
  
  # create vector with that pc for number of rows:
  pc_vec <- rep(pc_per_row, times = num_cols)
  
  
  # Loop 1: Simulations ---- 
  
  for (i in 1:n_sims){
    
    # Add NAs to data:
    data_NA <- replace_values(data = data, 
                              cols = col_names, 
                              amount = pc_vec, 
                              replacement = NA, 
                              levels_amount = NULL)
    
    # Apply tree definition to (OLD/NEW NA) data:
    y <- FFTrees(object = x,
                 tree.definitions = fft_df,
                 data = data_NA, 
                 quiet = TRUE, 
                 do.comp = FALSE)
    
    # Collect results:
    
    # a) tree performance: 
    y_summary <- summary(y)
    y_stats <- y_summary$stats$train
    
    # b) level stats:
    y_level_stats <- y$trees$level_stats$train
    NA_stats <- y_level_stats[nrow(y_level_stats), c(25:30)]
    
    # Combine:
    all_data_df <- cbind(y_stats, NA_stats)
    # dim(all_data_df)
    
    # Collect results:
    results[(1 + i), ] <- all_data_df
    
  } # for loop 1. 
  
  
  # After loop:
  
  # Add column names to data frame:
  names(results) <- names(all_data_df)
  
  
  # Output: ----
  return(results) # as data frame.
  
} # loop_NA().
```

## Run simulations

### Run 25% NA

```{r apply-loop-function-25-na}
# use loop-function, input = 25% nas:
heart_NA_25 <- loop_NA(data = heartdisease, 
                       criterion = "diagnosis", 
                       pc = 0.25)
```

### Results 25% NA

```{r results-25-na}
heart_NA_25
```

Collect key results (in df): 

```{r key-results-25}
key_results <- data.frame(matrix(NA, nrow = 5, ncol = 4))
names(key_results) <- c("pc_NA", "mn_acc", "mn_bacc", "mn_mcu")

# Baseline performance:
key_results[1, "pc_NA"]   <- 0
key_results[1, "mn_acc"]  <- heart_NA_25$acc[1]
key_results[1, "mn_bacc"] <- heart_NA_25$bacc[1]
key_results[1, "mn_mcu"]  <- heart_NA_25$mcu[1]

# Row 2: 25% NA values:
key_results[2, "pc_NA"]   <- 25
key_results[2, "mn_acc"]  <- mean(heart_NA_25$acc[-1])
key_results[2, "mn_bacc"] <- mean(heart_NA_25$bacc[-1])
key_results[2, "mn_mcu"]  <- mean(heart_NA_25$mcu[-1])

key_results
```

### Run 50% NA

```{r apply-loop-function-50-na}
# use loop-function, input = 50% nas:
heart_NA_50 <- loop_NA(heartdisease, "diagnosis", 0.50)
```

### Results 50% NA

```{r results-50-na}
heart_NA_50
```

Collect key results (in df): 

```{r key-results-50}
# key_results <- data.frame(matrix(NA, nrow = 4, ncol = 3))
# names(key_results) <- c("pc_NA", "mn_bacc", "mn_mcu")

# Row 3: 50% NA values:
key_results[3, "pc_NA"]   <- 50
key_results[3, "mn_acc"]  <- mean(heart_NA_50$acc[-1])
key_results[3, "mn_bacc"] <- mean(heart_NA_50$bacc[-1])
key_results[3, "mn_mcu"]  <- mean(heart_NA_50$mcu[-1])

key_results
```


### Run 75% NA

```{r apply-loop-function-75-na}
# use loop-function, input = 75% nas:
heart_NA_75 <- loop_NA(heartdisease, "diagnosis", 0.75)
```

### Results 75% NA

```{r results-75-na}
heart_NA_75
```

Collect key results (in df): 

```{r key-results-75}
# key_results <- data.frame(matrix(NA, nrow = 4, ncol = 3))
# names(key_results) <- c("pc_NA", "mn_bacc", "mn_mcu")

# Row 4: 75% NA values:
key_results[4, "pc_NA"]   <- 75
key_results[4, "mn_acc"]  <- mean(heart_NA_75$acc[-1])
key_results[4, "mn_bacc"] <- mean(heart_NA_75$bacc[-1])
key_results[4, "mn_mcu"]  <- mean(heart_NA_75$mcu[-1])

key_results
```



### Run 100% NA

Basically a sanity check: 

```{r apply-loop-function-100-na}
# how many rows are in the data set:
n <- nrow(heartdisease)

# use loop-function, input = 100% nas:
heart_NA_100 <- loop_NA(heartdisease, "diagnosis", n)
```

### Results 100% NA

```{r results-100-na}
heart_NA_100
```

Collect key results (in df): 

```{r key-results-100}
# key_results <- data.frame(matrix(NA, nrow = 4, ncol = 3))
# names(key_results) <- c("pc_NA", "mn_bacc", "mn_mcu")

# Row 4: 75% NA values:
key_results[5, "pc_NA"]   <- 100
key_results[5, "mn_acc"]  <- mean(heart_NA_100$acc[-1])
key_results[5, "mn_bacc"] <- mean(heart_NA_100$bacc[-1])
key_results[5, "mn_mcu"]  <- mean(heart_NA_100$mcu[-1])

key_results
```

Plot key results:

1. Accuracy (mean `acc`)

```{r plot-key-results-accuracy}
# Upper benchmark: Best FFT without missing data: 
plot(x = key_results$pc_NA, 
     y = rep(key_results$mn_acc[1], nrow(key_results)), 
     type = "l", 
     col = unikn::Seegruen, lwd = 1, lty = 2, 
     ylim = c(0, 1),
     xlab = "Percent NA values", 
     ylab = "Mean accuracy"
     )

grid()

# Lower benchmark pred_benchmark (i.e., majority of criterion):
lines(x = key_results$pc_NA, 
      y = rep(pred_benchmark, nrow(key_results)), 
      type = "l", 
      col = unikn::Bordeaux, lwd = 1, lty = 2 
      # ylim = c(0, 1),
      # xlab = "Percent NA values", 
      # ylab = "Mean accuracy"
)

# Actual accuracy:
lines(x = key_results$pc_NA, 
      y = key_results$mn_acc, 
      type = "b", 
      col = unikn::Seeblau, lwd = 2
      # ylim = c(0, 1),
      # xlab = "Percent NA values", 
      # ylab = "Mean accuracy"
)

title(main = "Mean accuracy (acc) given missing values")
```

2. Frugality (mean `mcu`)

```{r plot-key-results-frugality}

n_nodes <- nrow(read_fft_df(fft_df))

# Upper benchmark: Best FFT without missing data: 
plot(x = key_results$pc_NA, 
     y = rep(1, nrow(key_results)), 
     type = "l", 
     col = unikn::Seegruen, lwd = 1, lty = 2, 
     ylim = c(0, n_nodes + 1),
     xlab = "Percent NA values", 
     ylab = "Mean frugality (mcu)"
     )

grid()

# Lower benchmark pred_benchmark (i.e., majority of criterion):
lines(x = key_results$pc_NA, 
      y = rep(n_nodes, nrow(key_results)), 
      type = "l", 
      col = unikn::Bordeaux, lwd = 1, lty = 2 
      # ylim = c(0, 1),
      # xlab = "Percent NA values", 
      # ylab = "Mean accuracy"
)

# Actual frugality:
lines(x = key_results$pc_NA, 
      y = key_results$mn_mcu, 
      type = "b", 
      col = unikn::Seeblau, lwd = 2
      # ylim = c(0, 1),
      # xlab = "Percent NA values", 
      # ylab = "Mean accuracy"
)

title(main = "Mean frugality given missing values")
```




## Issues and potential problems 

+++ here now +++ 


### 1. Using <NA> in categorical predictors

- When creating the basic FFT _without_ any `NA` values, 
the mechanism for _categorical_ predictors (i.e., treating `NA` as a separate `<NA>` category) is not being used or tested.  
Thus, we should create the initial FFT (and corresponding performance baselines) for datasets with _some_ `NA` values (e.g., 10%).


### 2. Dealing with NA in final predictors

- The current default for coping with `NA` values in final node 
appears to be to assign all cases 
to the _majority_ class (as only 2 types of errors occur): 
`fin_NA_pred = "majority"` (in **FFTrees** v2.0.0). 

- Done: Switched default to `fin_NA_pred = "baseline"` in `fftrees_apply()` (i.e., biased coin flip with both outcomes).


### 3. Value of mcu?

- Why is the final `mcu` value (for simulations with 100% NA values) 
a value of\ 2, rather than\ 3 (for FFT `fft_df` with 3\ nodes)? 



<!-- eof. -->
